### 2010-03-22 thesis/SensorMLOnSenseTile/trunk/thesis/SensorMLonSenseTileThesis.tex
latest

\section{Sensor Modeling Language}
Sensor Modeling Language (SensorML) is primarily used to describe sensor systems and the processing of observations from sensor systems. The SensorML specification\cite{SMLref} provides a functional model of sensors and an XML encoding to describe sensors and their observations. Some of the main uses of SensorML described in the specification are:
\begin{itemize}
\item support sensor and sensor observation discovery
\item support processing of the sensor observations
\item support geolocation information about sensors
\item provide accuracy information of sensors measurements
\item Provide an explicit description of the process by which an observation was obtained
\item Provide an executable process chain for deriving new data products on demand
\end{itemize}

In SensorML sensors are modeled as processes that convert real phenomena to data. Processes take inputs, process them and result in one or several outputs. For example, a process can take a digital number generated by a thermistor and transform it to a celcius value. Processes can also be connected together in process chains to allow the sensor data to be processed by several processes sequentially. These processes and process chains are modeled in SensorML as shown in the  below SensorML Conceptual Model \ref{fig:SMLConceptualModel}.

@@

It is stated in the SensorML specification that the SWE Common will be placed in it's own document in future releases.

@@

\section{Sensor Observation Service}
SensorML is envisioned to be used in conjunction with a Sensor Observation Service as described in the SWE Architecture \cite{SWEArchref}. The SOS defines a web service interface for the discovery of sensors systems and the retrieval of observations from sensor systems. The SenseTile Web Service developed in this thesis is based on the concepts described in the Sensor Observation Service (SOS) specification.

The SOS specification uses four main entities to describe the SOS's operation. The entities are Data Consumers, Data Producers, OGC Catalogs and the SOS itself. Data Consumers are clients of the SOS that request sensor observations and sensor metadata. Data Providers are entities have to access to the sensor and have enough processing power to generate an XML encoding of the sensor's observations. These observations are sent over a network to the SOS. The OGC Catalog is used store metadata about the sensor systems and is used as part of the discovery process.

The diagram \ref{fig:SOSoperationContext} below shows the entities described in the SOS in an operation context
\begin{figure}[h]
\framebox{\scalebox{0.65}{\includegraphics{SOSoperationContext.png}}}
\caption{SOS Operational Context}\label{fig:SOSoperationContext}
\end{figure}

Collections of related sensor system observations are organized into Observation
Offerings.

Observations are encoded as OGC O\&M Observations. Sensor systems metadata can be also be retrieved using the SOS interface and is generally defined in SensorML or TML.
para.  These operations are encoded in XML and used in http post and get operations. The XML message needs to be parsed to identify the operations, parameters and results. There seems to be work ongoing to change this to a more industry standard view of Web Services with the use of WSDL/SOAP. I could not find this as part of current OGC SWE framework.
para: Scalability would be achieved through aggregating a number of
sensor data producers on the sensor side and by aggregating a number of SOS instances
on the data center side for providing data to consumers. 
\subsection{SOS Data Consumer Operations}
There are there core operations that are provided by an SOS and used by SOS Clients to request sensor information:
 \begin{itemize}
\item GetObservation operation is used to obtain sensor observations and measurement data.
\item GetCapabilities operation is used to retrieve SOS service metadata.
\item DescribeSensor operation retrieves detailed information about the sensors.
\end{itemize}

\subsection{SOS Data Provider Operations}
Two other operations RegisterSensor and InsertObservation are of interest to this thesis. These are part of the SOS transactional interface. They are used with SOS data producers which are software entities connected to the sensor and have enough processing power to generate XML descriptions of the sensor data.

\subsection{SOS Catalogs}
The catalog aspect of SOS that used for discovery of the SOS instance is not covered in the Web Service Prototype described in this thesis. 
\subsection{Enhanced Operations}
Also six enhanced operations, GetResult,
GetFeatureOfInterest, GetFeatureOfInterestTime, DescribeFeatureOfInterest,
DescribeObservationType, and DescribeResultModel are not included in the SenseTileWeb Service.

@@

This SenseTile unit is where thesis will focus on the use of SensorML though SensorML can be used in other the parts of the SenseTile system. As the processor unit is  a relatively powerful computer, providing a Web Service on the SenseTile Unit is feasible. The part SensorML plays in this Web Service is explored in the context of the SenseTile system.

@@

\section{VAST SensorML Engine}
The VAST Team at the University of Alabama in Huntsville (UAH)  developed an open source JAVA based SensorML Processing Engine. This software parses SensorML and instantiates objects that implement SensorML processes and process  chains. The code is found at http://code.google.com/p/sensorml-data-processing/. It dependent on an implementation of the SWE Common data types described in the chapter on SensorML. This code can be found at call swe-common-data-framework http://code.google.com/p/swe-common-data-framework/. This software is referred to as the VastSMLEngine in rest of the thesis.

@@

@@

\chapter{ Conclusions and Future Work}

what has been achieved

@@

\chapter{References}
\newpage
\begin{thebibliography}{99}
\bibitem{SMLref}Open Geospatial Consortium Inc., OpenGIS Sensor Model Language (SensorML) Implementation Specification, 2007
\bibitem{SOSref}Open Geospatial Consortium Inc.,  Sensor Observation Service, 2007
\bibitem{OMref}Open Geospatial Consortium Inc., Observations and Measurements – Part 1 - Observation schema, 2007
\bibitem{SWEArchref}Open Geospatial Consortium Inc., OGC Sensor Web Enablement Architecture, 2008
\bibitem{SMLTutorialref}Open Geospatial Consortium Inc., Using SensorML to describe a
Complete Weather Station , 2006
\bibitem{ProcessTutorialref}Open Geospatial Consortium Inc.,

\bibitem{BONref}Kim Waldén and Jean-Marc Nerson , "Seamless Object-Oriented Software Architecture", 1995
\end{thebibliography}
\label{endpage}

>>>>

\section{Sensor Modeling Language}\label{SMLsection}
Sensor Modeling Language (SensorML) is primarily used to describe sensor systems and the processing of observations from sensor systems. The SensorML specification\cite{SMLref} provides a functional model of sensors and an XML encoding to describe sensors and their observations. Some of the main uses of SensorML described in the specification are:
\begin{itemize}
\item support sensor and sensor observation discovery
\item support processing of the sensor observations
\item support geolocation information about sensors
\item provide accuracy information of sensors measurements
\item Provide an explicit description of the process by which an observation was obtained
\item Provide an executable process chain for deriving new data products on demand
\end{itemize}

In SensorML sensors are modeled as processes that convert real phenomena to data. Processes take inputs, process them and result in one or several outputs. For example, a process can take a measurement generated by a thermistor which might be a resistance value and transform it to a celcius value. Processes can also be connected together in process chains to allow the sensor data to be processed by several processes sequentially. These processes and process chains are modeled in SensorML as shown in the  below SensorML Conceptual Model \ref{fig:SMLConceptualModel}.

@@

SensorML makes extensive use of XLink components to support hypertext referencing in XML. This allows one to reuse Objects that are either internal or external to the instance document, and allows these objects to perhaps have different associations or roles within one or more documents. This is supported by extensive use of the id attribute (taking an xs:ID as its value) within Objects, and utilizing the GML attribute group, gml:AssociationGroup, within property elements.
SensorML is designed with the concept that there exist dictionaries or registries for definitive specification of various terms.
these definitions should be specified in SensorML documents as resolvable URNs, while XLink references will be either URLs or URNs.

@@

\section{Sensor Observation Service}
SensorML is envisioned to be used in conjunction with a Sensor Observation Service as described in the SWE Architecture \cite{SOSref}. The SOS defines a web service interface for the discovery of sensors systems and the retrieval of observations from sensor systems.  The SenseTile Web Service developed in this thesis is based on the concepts described in the Sensor Observation Service (SOS) specification.

The SOS specification uses four main entities to describe the SOS's operation. The entities are Data Consumers, Data Producers, OGC Catalogs and the SOS itself. Data Consumers are clients of the SOS that request sensor observations and sensor metadata. Data Providers are entities have to access to the sensor and have enough processing power to generate an XML encoding of the sensor's observations. These observations are sent over a network to the SOS. The OGC Catalog is used store metadata about the sensor systems and is used as part of the discovery process. The SOS is an access point to a Sensor Web Network and provides the Web Service interface. The figure \ref{fig:SOSoperationContext} below shows the entities described in the SOS in an operation context.
\begin{figure}[h]
\framebox{\scalebox{0.65}{\includegraphics{SOSoperationContext.png}}}
\caption{SOS Operational Context}\label{fig:SOSoperationContext}
\end{figure}

The diagram above shows the core mandatory operations and transactional operations provided by an SOS. The core operations are used by Data Consumers to request sensor information. The operations perfom the following function:
 \begin{itemize}
\item GetObservation operation is used to obtain sensor observations and measurement data.
\item GetCapabilities operation is used to retrieve SOS service metadata.
\item DescribeSensor operation retrieves detailed information about the sensors.
\end{itemize}

The transactional operations supported by the SOS toward the DataProvider are:
 \begin{itemize}
\item RegisterSensor operation registers the DataProvider with the SOS.
\item InsertObservation operation updates SOS service metadata.
\end{itemize}

Collections of related sensor system observations are organized by the SOS into observation offerings. Observations are encoded as OGC O\&M Observations.

These operations are encoded in XML and used in http post and get operations. The XML message needs to be parsed to identify the operations, parameters and results.

Sensor systems metadata can be also be retrieved using the SOS interface and is generally defined in SensorML or TML.
para.

The SOS Operation Context diargam also shows a discovery process involving the  Data Consumer and a OGC Catalog Service (CS-W) catalog. The Data Consumer uses a CS-W catalog to find SOS instances. The GetCapabilites are then used toward each of the SOS instances returned to find sensor systems that are of interest to the Data Consumer. The CS-W catalog is covered seperate specification \cite{OGCcatref} and is not covered in any real detail in the SOS specification. The OGC Catalog Service is not explored in this thesis.

There are also six other operations the SOS refers to as enhanced operations. They are GetResult,
GetFeatureOfInterest, GetFeatureOfInterestTime, DescribeFeatureOfInterest,
DescribeObservationType, and DescribeResultModel. These operations are not covered in this thesis.

@@

This SenseTile unit is used here as case study on the use of SensorML, though SensorML can be used in other the parts of the SenseTile system. As the processor unit is  a relatively powerful computer, providing a Web Service on the SenseTile Unit is feasible. The part SensorML plays in this Web Service is explored in the context of the SenseTile system.

Thermistor?

@@

\section{VAST SensorML Engine}
The VAST Team at the University of Alabama in Huntsville (UAH)  developed an open source Java based SensorML Processing Engine. This software provides types that implement SensorML processes and process  chains. The SensorML Processing Engine parses SensorML models and instantiates the required objects to performed the intended processing. The code is found at http://code.google.com/p/sensorml-data-processing and is referred to as the VastSMLEngine in rest of the thesis. This code is dependent on an implementation of the SWE Common data types described in the section on SensorML \ref{SMLsection}. This code is found at http://code.google.com/p/swe-common-data-framework.

@@

Scalability - info here.

@@

\chapter{ Conclusions and Future Work}
There seems to be work ongoing to change this to a more industry standard view of Web Services with the use of WSDL/SOAP. I could not find this as part of current OGC SWE framework.
what has been achieved

@@

\chapter{References}
\newpage
\begin{thebibliography}{99}
\bibitem{SMLref}Open Geospatial Consortium Inc., OpenGIS Sensor Model Language (SensorML) Implementation Specification, 2007
\bibitem{SOSref}Open Geospatial Consortium Inc.,  Sensor Observation Service, 2007
\bibitem{OMref}Open Geospatial Consortium Inc., Observations and Measurements – Part 1 - Observation schema, 2007
\bibitem{SWEArchref}Open Geospatial Consortium Inc., OGC Sensor Web Enablement Architecture, 2008
\bibitem{SMLTutorialref}Open Geospatial Consortium Inc., Using SensorML to describe a
Complete Weather Station , 2006
\bibitem{ProcessTutorialref}Open Geospatial Consortium Inc.,
\bibitem{OGCcatref}Open Geospatial Consortium Inc., OpenGIS® Catalogue Services Specification, 2007
\bibitem{BONref}Kim Waldén and Jean-Marc Nerson , "Seamless Object-Oriented Software Architecture", 1995
\end{thebibliography}
\label{endpage}
